Re-analysing the data from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}): A textbook illustration of the absence of evidence fallacy

### 2020-10-07 manuscript.Rmd
updating the bf dance simulation

```{r bf-dance, fig.cap = "Illustrating the distribution of Bayes factors (also known as the 'Dance of the Bayes factors') in favour of the alternative hypothesis for different effect sizes (N = 26). In the left panel, the effect size is fixed to d = 0 (i.e., the null hypothesis), in the middle panel, it is fixed to d = 0.36 (i.e., the supposed target effect size in Moffatt et al., 2020), and in the right panel, the effect size is fixed to d = 0.72 (i.e., the effect size reported in Nalborcayk et al., 2017)."}
# https://julianquandt.com/post/power-analysis-by-data-simulation-in-r-part-ii/
# http://daniellakens.blogspot.com/2016/07/dance-of-bayes-factors.html
# population effect size
cohensd <- 0.36

@@

# number of simulated experiments
nsims <- 1e3

@@

First, we simulated 10.000 datasets (for N = 26) under the assumption of either no effect (i.e., the null hypothesis), an effect size of d = 0.36 [i.e., the supposed target effect size in @moffatt_inner_2020], or an effect size of d = 0.72 [i.e., the effect size reported in @nalborczyk_orofacial_2017]. As shown in Figure \@ref(fig:bf-dance), the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis (i.e., when the "true" effect size is d = 0), `r round(prop_d0 * 100, 2)`% of the computed BFs are above 0 and hence support the alternative hypothesis. When the "true" effect size is of d = 0.36, `r round(100 - prop_d0.36 * 100, 2)`% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the "true" effect size is of d = 0.72, `r round(100 - prop_d0.72 * 100, 2)`% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.

>>>>

```{r bf-dance, fig.cap = "Illustrating the distribution of Bayes factors (also known as the 'Dance of the Bayes factors') in favour of the alternative hypothesis for different effect sizes (N = 26). In the left panel, the effect size is fixed to d = 0 (i.e., the null hypothesis), in the middle panel, it is fixed to d = 0.36 (i.e., the supposed target effect size in Moffatt et al., 2020), and in the right panel, the effect size is fixed to d = 0.72 (i.e., the effect size reported in Nalborczyk et al., 2017)."}
# https://julianquandt.com/post/power-analysis-by-data-simulation-in-r-part-ii/
# http://daniellakens.blogspot.com/2016/07/dance-of-bayes-factors.html
# population effect size
cohensd <- 0.36

@@

# number of simulated experiments
nsims <- 1e4

@@

First, we simulated 10.000 datasets (for N = 26) under the assumption of either no effect (i.e., the null hypothesis of d = 0), an effect size of d = 0.36 [i.e., the supposed target effect size in @moffatt_inner_2020], or an effect size of d = 0.72 [i.e., the effect size reported in @nalborczyk_orofacial_2017]. As shown in Figure \@ref(fig:bf-dance), the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis, `r round(prop_d0 * 100, 2)`% of the computed BFs are above 0 and hence support the alternative hypothesis (although the "true" effect size is d = 0). When the "true" effect size is of d = 0.36, `r round(100 - prop_d0.36 * 100, 2)`% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the "true" effect size is of d = 0.72, `r round(100 - prop_d0.72 * 100, 2)`% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.^[To assess the extent to which the BF10 computed for the OOI by @moffatt_inner_2020 is "surprising" given or "compatible" with the sampling distribution of BFs under the hypothesis of an effect size of d = 0.36, we can compute the probability of obtaining this finding or a more extreme finding given the hypothesis, which is approximately equal to `r round(1 - mean(results$BF10 > 0.278), 2)`.]

### 2020-10-07 manuscript.tex
updating the bf dance simulation

Wordcount (excluding abstract, references, tables, and figures): 4306

@@

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the \emph{relative evidence} for the null, computing a BF (i.e., comparing two models) does not solve at all the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this, we present below the results of a simulation aiming to assess the relation between the sample size and the value of the BF.

In the previous section, we fitted a multivariate Bayesian regression model with varying-intercepts by participant and weakly informative priors on the EMG data collected by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}). Using this model, we i) generated new datasets from the posterior predictive distribution of this model and ii) we computed the BF in favour of the alternative hypothesis (\(\text{BF}_{10}\)) using the \texttt{BayesFactor} package (Morey \& Rouder, \protect\hyperlink{ref-R-BayesFactor}{2018}). We used a ``medium'' prior (i.e., a scale of 1) on the scale of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with \(1000\) simulations (i.e., 1000 simulated datasets) for each sample size.

>>>>

Wordcount (excluding abstract, references, tables, and figures): 4523

@@

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the \emph{relative evidence} for the null, computing a BF (i.e., comparing two models) does not solve at all the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this point, we present below the results of two simulations.

{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/bf-dance-1}

\caption{Illustrating the distribution of Bayes factors (also known as the 'Dance of the Bayes factors') in favour of the alternative hypothesis for different effect sizes (N = 26). In the left panel, the effect size is fixed to d = 0 (i.e., the null hypothesis), in the middle panel, it is fixed to d = 0.36 (i.e., the supposed target effect size in Moffatt et al., 2020), and in the right panel, the effect size is fixed to d = 0.72 (i.e., the effect size reported in Nalborczyk et al., 2017).}\label{fig:bf-dance}
\end{figure}

First, we simulated 10.000 datasets (for N = 26) under the assumption of either no effect (i.e., the null hypothesis of d = 0), an effect size of d = 0.36 (i.e., the supposed target effect size in Moffatt et al., \protect\hyperlink{ref-moffatt_inner_2020}{2020}), or an effect size of d = 0.72 (i.e., the effect size reported in Nalborczyk et al., \protect\hyperlink{ref-nalborczyk_orofacial_2017}{2017}). As shown in Figure \ref{fig:bf-dance}, the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis, 6.96\% of the computed BFs are above 0 and hence support the alternative hypothesis (although the ``true'' effect size is d = 0). When the ``true'' effect size is of d = 0.36, 71.83\% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the ``true'' effect size is of d = 0.72, 25.58\% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.\footnote{To assess the extent to which the BF10 computed for the OOI by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) is ``surprising'' given or ``compatible'' with the sampling distribution of BFs under the hypothesis of an effect size of d = 0.36, we can compute the probability of obtaining this finding or a more extreme finding given the hypothesis, which is approximately equal to 0.29.}

Second, we conducted another simulation with the aim of assessing the relation between the sample size and the value of the BF. In the previous section, we fitted a multivariate Bayesian regression model with varying-intercepts by participant and weakly informative priors on the EMG data collected by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}). Using this model, we i) generated new datasets from the posterior predictive distribution of this model and ii) we computed the BF in favour of the alternative hypothesis (\(\text{BF}_{10}\)) using the \texttt{BayesFactor} package (Morey \& Rouder, \protect\hyperlink{ref-R-BayesFactor}{2018}). We used a ``medium'' prior (i.e., a scale of 1) on the scale of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with \(1000\) simulations (i.e., 1000 simulated datasets) for each sample size.
